### 2022-04-18 Draft2.Rmd
The data is loaded in and I added some questions to consider

---
title: "Draft2"
output: slidy_presentation
date: '2022-04-18'
---

# Load libraries 
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Add in Data 
```{r}
read_csv('death-rates-from-air-pollution.csv')
```

# Questions to Consider

## Which country has the greatest number of deaths? The least?

## Which type of air pollution has the greatest number of deaths (indoor, outdoor, particulate, or outdoor ozone)?

## Which year had the worst air pollution (average the three diff types)?

## Which year had the worst indoor? Outdoor particulate? Outdoor ozone?