### 2022-04-18 Draft2.Rmd
made a table for increments of 5 years from 2007-2017. having issues with the graph

---
title: "Draft2"
output: slidy_presentation
date: '2022-04-18'
---

# Load libraries 
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Add in Data 
```{r}
deaths_df <- data.frame(read.csv("death-rates-from-air-pollution.csv"))
deaths_df

colnames(deaths_df) <- c("country", "acronym", "year", "total_deaths", 
                         "indoor_deaths", "outdoor_deaths", "ozone_deaths")

glimpse(deaths_df)
```

# Questions to Consider

## Which country has the greatest number of deaths? The least? (dependent on population)

```{r}
# Want the range 2007-2017 
# took this out c(2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017)

#Maybe try increments...?
#When looking at it you can see a decrease among the three years 
death_07<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2007) %>% 
  select(total_deaths) %>% 
  summarize(average_death_07 = mean(total_deaths))

death_12<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2012) %>% 
  select(total_deaths) %>% 
  summarize(average_death_12 = mean(total_deaths))

death_17<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2017) %>% 
  select(total_deaths) %>% 
  summarize(average_death_17 = mean(total_deaths))

#try to have one cohesive table so it's easier to visualize 
combo <- death_07 %>% 
  mutate(death_12) %>% 
  mutate(death_17)

combo

#Then graph it 
#Just realized there's 200+ countries....the x-axis looks terrible 
#Would be nice to color by continent but then that would mean going through each country
ggplot(death_07, aes(x= country, y=average_death_07)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 15)) #yah..no rotation made it worse

#Not sure if we want to use the selected few countries we looked at earlier?
#Or if we wanted to use this as a overview of the data
```

## Which type of air pollution has the greatest number of deaths (indoor, outdoor, particulate, or outdoor ozone)?

```{r}

```

## Which year had the worst air pollution (average the three diff types)?

```{r}

````

## Which year had the worst indoor? Outdoor particulate? Outdoor ozone?

```{r}

````

# Notes from first meeting

## Hone in to only one year. Find data that is only population, continent data. Choose a couple countries/continents to work on

## Look at industrialized vs non-industrialized countries