### 2022-04-18 Project_draft.Rmd

---
title: "Project draft"
author: "Alison Cheek"
date: "4/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggplot2)

```

#Import air pollution deaths csv and tidy

```{r}

deaths_df <- data.frame(read.csv("death-rates-from-air-pollution.csv", stringsAsFactors = FALSE))

glimpse(deaths_df)

colnames(deaths_df)

colnames(deaths_df) <- c("country", "acronym", "year", "total_deaths", 
                         "indoor_deaths", "outdoor_deaths", "ozone_deaths")

```

#Explore data
```{r}
#glimpse(deaths_df)

deaths_df %>% filter(country=="United States") %>%
  select(year, indoor_deaths) %>% 
  ggplot() + 
  geom_point(aes(x=year, y=indoor_deaths))

deaths_df %>% filter(country=="United States") %>%
  select(year, outdoor_deaths) %>% 
  ggplot() + 
  geom_point(aes(x=year, y=outdoor_deaths))

deaths_df %>% filter(country=="United States") %>%
  select(year, ozone_deaths) %>% 
  ggplot() + 
  geom_point(aes(x=year, y=ozone_deaths))
```

#
```{r}
deaths_df %>% 
  select(total_deaths, indoor_deaths, outdoor_deaths, ozone_deaths) %>% 
  filter(deaths_df$acronym == "AFG") %>% 
  summarize(mean(total_deaths))

deaths_df %>% 
  select(total_deaths, indoor_deaths, outdoor_deaths, ozone_deaths) %>% 
  filter(deaths_df$acronym == "USA") %>% 
  summarize(mean(total_deaths))
```

#

#

#

#

#

#