### 2022-04-25 population_data.Rmd
population data

##Population data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggplot2)

```

```{r}
pop_data_stats <- read.csv("pop_worldometer_data.csv")

glimpse(pop_data_stats)

world_pop <- read.csv("population_total_long.csv")

glimpse(world_pop)

high_pop_countries <- world_pop %>% 
  filter(Year > 1999 & Country.Name %in% c('United States', 'Brazil', 'Ethiopia', 'Russia', 'China', 'Australia')) %>% 
  group_by(Year)

high_pop_countries

low_pop_countries <- world_pop %>% 
  filter(Year > 1999 & Country.Name %in% c('United States', 'Brazil', 'Ethiopia', 'Russia', 'China', 'Australia'))%>% 
  group_by(Year)

low_pop_countries

```