### 2022-04-25 population_data.Rmd
changes were made after second meeting

##Population data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggplot2)

```

```{r}
pop_data_stats <- read.csv("pop_worldometer_data.csv")

glimpse(pop_data_stats)

world_pop <- read.csv("population_total_long.csv")

glimpse(world_pop)
head(world_pop)

```

The above is 1960/2020 and our data is from 1990-2017
This is a different file but won't open
```{r}
#This messed it up into only 1 column sep ='\t', ;, 
pops_by_country <- read.csv("world_population_update.csv")
head(pops_by_country)

#
```