### 2022-04-29 population_data.Rmd
These is the population data with countries and years we want.

##Population data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggplot2)

```

```{r}
pop_data_stats <- read.csv("pop_worldometer_data.csv")

glimpse(pop_data_stats)

world_pop <- read.csv("population_total_long.csv")

glimpse(world_pop)
head(world_pop)

```

The above is 1960/2020 and our data is from 1990-2017
This is a different file but won't open (resolved)
```{r}
#This is the data 
pops_by_country <- read.csv("world_population_update.csv")
head(pops_by_country)

#The columns that we want 
pops_by_country %>% filter(Country.Name %in% c('United States', 'Brazil', 'Ethiopia','Russian Federation','China', 'Australia','Canada','Argentina','Seychelles','Romania','Israel', 'Tonga')) %>% select(Country.Name, Country.Code, X1990, X1991, X1992, X1993, X1994, X1995, X1996, X1997, X1998, X1999, X2000, X2001, X2002, X2003, X2004, X2005, X2006, X2007, X2008, X2009, X2010, X2011, X2012, X2013, X2014, X2015, X2016, X2017)
```