### 2022-05-02 outline.Rmd
updates

---
title: "Global Deaths Due to Air Pollution"
author: "Elizabeth Bekele, Alison Cheek"
output: slidy_presentation
date: '2022-05-03'
---

# Introduction 
* Air pollution can be detrimental to both our health and the climate
  + Outdoor and indoor air pollution cause chronic pain, respiratory diseases, shortened lifespan
  + Air pollution kills about 7 million people worldwide every year 
  + Hopefully this information can showcase the importance of air pollution and that we should be more mindful about our planet

* Overview
  + We will see how different air pollution types affect the population 
  + compare past and present population numbers
  + determine which air pollutant type has the highest associated death rate

# Packages Required 
```{r, message=FALSE, warning=FALSE}
#This will allow us to filter through our data 
library(tidyverse)
library(dplyr)
#This will help us plot figures to showcase our findings
library(ggplot2)
#This will help us organize and display our data as necessary 
library(knitr)
library(kableExtra)
#This expands our plot uses 
library(plotly)
#Scientific Notation Disabled 
options(scipen=999)
```

# Deaths Data
* Deaths from pollution:
  + Kaggle: [https://www.kaggle.com/datasets/akshat0giri/death-due-to-air-pollution-19902017] 
  + Author: Akshat Giri
  + Last update: 2 years ago

Import the deaths-due-to-air-pollution data
```{r}
deaths_df <- data.frame(read.csv("death-rates-from-air-pollution.csv"))
```

We are going to rename a few of the columns and glimpse the data
```{r}
colnames(deaths_df) <- c("country", "acronym", "year", "total_deaths", "indoor_deaths", "outdoor_deaths", "ozone_deaths")

glimpse(deaths_df)
```

# Data Variables
Variables that interest us here include:

* __country__ 
* __total_deaths__: per 100,000
* __indoor_deaths__: Indoor air pollution is considered pollution that occurs in the household. Cooking with solid fuels:
  + Wood
  + Crop waste, dung
  + Charcoal, coal
* __outdoor_deaths__: Outdoor air or ambient air are emissions caused by combustion processes from motor vehicles, solid fuel burning and industry 
  + Ozone (O3)
  + Particulate matter (PM10 and PM2.5)
  + Nitrogen dioxide (NO2)
  + Carbon monoxide (CO)
  + Sulfur dioxide (SO2)
* __ozone_deaths__: Ozone is a gas that occurs both in Earth's upper atmosphere and at ground level. Ozone in the atmosphere is an important and helpful greenhouse gas, but ground-level ozone is created by extensive use of fossil fuels: 
  + Pollutants emitted by cars
  + Power plants, industrial boilers, refineries, chemical plants

# World Population Data 
* World Population:
  + Kaggle: [https://www.kaggle.com/datasets/imdevskp/world-population-19602018] 
  + Author: Devakumar K. P. 
  + Last update: 2 Years ago

Now, let's take a look at the population data.
```{r}
world_pop <- read.csv("population_total_long.csv")
glimpse(world_pop)
```

***

To get a general idea of 'deaths-dataframe' we made, let's make a plots to see what's happening. This is a plot of indoor x outdoor deaths around the world by country.

```{r, echo=FALSE}
d <- ggplot(deaths_df, aes(x = indoor_deaths, y = outdoor_deaths, text = paste0(country, ", ", year) )) + geom_point() +
  ggtitle("Outdoor Deaths vs Indoor Deaths")
ggplotly(d)
```

This is a mess, and so we chose two countries from each continent (a high-population and a low-population country) to graph.

***
We selected a high population from each continent and used the formula below to determine the low population.

Low population = high population * .10

```{r, echo=FALSE}
#selecting high-population countries from the world population data frame 
high_pop_countries <- world_pop %>% 
  filter(Year > 1996 & Country.Name %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(Year)

ggplot(high_pop_countries, aes(x=Country.Name, y=Count, color=Country.Name))+
  geom_boxplot()+
  labs(title="Selection of high-population countries", x="Country", y="Population (in millions)")

#selecting low-population countries from the world population data frame 
low_pop_countries <- world_pop %>% 
  filter(Year > 1996 & Country.Name %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand'))%>% 
  group_by(Year)

ggplot(low_pop_countries, aes(x=Country.Name, y=Count, color=Country.Name))+
  geom_boxplot()+
  labs(title="Selection of low-population countries", x="Country", y="Population (in millions)")

kable(head(list(high_pop_countries[1:6,], low_pop_countries[1:6,])))

```

# Combine Data Sets

First let's look at a table of the high and low populated countries using the world population data set.
```{r, echo=FALSE}
kable(head(list(high_pop_countries[1:6,], low_pop_countries[1:6,])))
```

Next, we are going to see the death count for high and low populated countries using the deaths dataframe.
```{r, echo=FALSE}
#selecting high-population deaths from death dataframe 
high_pop_death <- deaths_df %>% 
  filter(year > 1996 & country %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(year)

#selecting low-population deaths from death dataframe 
low_pop_death <- deaths_df %>% 
  filter(year > 1996 & country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand')) %>% 
  group_by(year)

kable(head(list(high_pop_death[1:22,], low_pop_death[1:22,])))
```

Lastly, we will join the population and and deaths with its respected country.
```{r, message=FALSE, echo=FALSE}
#Combined High Population with Pollution Death
joined_high <- right_join(high_pop_death, high_pop_countries, by= c('country' = 'Country.Name', 'year' = 'Year'))

#Combined Low Population with Pollution Death
joined_low <-right_join(low_pop_death, low_pop_countries, by= c('country' = 'Country.Name', 'year' = 'Year'))

kable(head(list(joined_high[1:22,], joined_low[1:22,])))
```

***

```{r, echo=FALSE}
ggplot(data=joined_high, aes(x=year, y=Count, color=country, size = total_deaths/Count)) +
  geom_line() +
  labs(title="Percentage of Death Count in High Population Countries", subtitle="Bar width is equal to percentage of Deaths")+
  xlab("Year")+
  ylab("Population")

ggplot(data=joined_low, aes(x=year, y=Count, color=country, size = total_deaths/Count)) +
  geom_line() +
  labs(title="Percentage of Death Count in Low Population Countries", subtitle="Bar width is equal to percentage of Deaths")+
  xlab("Year")+
  ylab("Population")
```

# Death Count
> Which country has the highest death count?

Let's make a table depicting the high and low populated countries and their respected death count due to pollution.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Mean total deaths of high-population countries
deaths_highpop_countries <- deaths_df %>% 
  filter(country %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country) %>% 
  select(total_deaths) %>% 
  summarize(hp_average_death = mean(total_deaths))

#Mean total deaths of high-population countries
deaths_lowpop_countries<- deaths_df %>% 
  filter(country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand')) %>% 
  group_by(country) %>% 
  select(total_deaths) %>% 
  summarize(lp_average_death = mean(total_deaths))

#Side-by-side Table to view deaths 
kable(list(deaths_highpop_countries, deaths_lowpop_countries))

```

***

Here's a graph to clearly visualize the previous table
```{r, echo=FALSE, message=FALSE}
#Plot High Population Deaths (average)
h<- ggplot(deaths_highpop_countries)+
  geom_col(mapping = aes(x=country, y=hp_average_death, color=country))+
             xlab("Country")+
             ylab("Average deaths (per 100,000)")+
             ggtitle("Average deaths in high-population countries")+
  coord_flip()

ggplotly(h)

#Plot of Low Population Deaths (average)
l <- ggplot(deaths_lowpop_countries)+
  geom_col(mapping = aes(x=country, y=lp_average_death, color=country))+
             xlab("Country")+
             ylab("Average deaths (per 100,000)")+
             ggtitle("Average deaths in low-population countries")+
  coord_flip()

ggplotly(l)
```

*** 
So we've looked at the deaths due to pollution, but what percentage of the population was affected?

```{r, echo=FALSE, message=FALSE}
hp_countries_population <- world_pop %>% 
  filter(Country.Name %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia'), Year > 1996) %>% 
  group_by(Country.Name) %>% 
  select(Count) %>% 
  summarize(average_population = mean(Count))

lp_countries_population <- world_pop %>% 
  filter(Country.Name %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand'), Year > 1996) %>% 
  group_by(Country.Name) %>% 
  select(Count) %>% 
  summarize(average_population = mean(Count))

#Population Average Table
kable(list(hp_countries_population, lp_countries_population))

#Percentage of those affected 
affected_high<- joined_high %>% 
  mutate(percent_high = total_deaths/mean(Count))

percent_high<- affected_high %>% ggplot(aes(x = year, y = percent_high, color = country))+
  geom_point() +
  labs(title = "High Population Countries Affected by Air Pollution") +
  xlab("Year")+
  ylab("Percent (total_deaths/average pop)")
ggplotly(percent_high)

affected_low <- joined_low %>% 
  mutate(percent_low = total_deaths/mean(Count))

percent_low <- affected_low %>% ggplot(aes(x = year, y = percent_low, color = country))+
  geom_point() +
  labs(title = "Low Populated Countries Affected by Air Pollution") +
  xlab("Year")+
  ylab("Percent (total_deaths/average pop)")
ggplotly(percent_low)
```

# Pollution Types
> Which type of pollution has the greatest number of deaths?

```{r, echo=FALSE, message=FALSE}
#High Population Pollutant Averages
high_poll <- deaths_df %>% 
  group_by(country) %>% 
  filter(country %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  select(country, indoor_deaths, outdoor_deaths, ozone_deaths) %>% 
  summarize(avg_indoor = mean(indoor_deaths), avg_outdoor = mean(outdoor_deaths), avg_ozone = mean(ozone_deaths)) %>% 
  arrange(desc(avg_indoor),desc(avg_outdoor),desc(avg_ozone))

kable(high_poll)

#Low Population Pollutant Averages 
low_poll <- deaths_df %>% 
  group_by(country) %>% 
  filter(country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand')) %>% 
  select(country, indoor_deaths, outdoor_deaths, ozone_deaths) %>% 
  summarize(avg_indoor = mean(indoor_deaths), avg_outdoor = mean(outdoor_deaths), avg_ozone = mean(ozone_deaths))
kable(low_poll)
```

***
```{r, echo=FALSE}

#Plot High Population
# High Indoor Air Pollution
h_indoor<- ggplot(high_poll, aes(x=country, y = avg_indoor, cex=5)) + 
  geom_point() +
  ggtitle("Indoor Air Pollution Deaths in High Population Countries")+
  xlab("Country")+
  ylab("Average Deaths")
ggplotly(h_indoor)

#Low Indoor Air Pollution
l_indoor<- ggplot(low_poll, aes(x=country, y = avg_indoor, cex=5)) + 
  geom_point() +
  ggtitle("Indoor Air Pollution Deaths in Low Population Countries")+
  xlab("Country")+
  ylab("Average Deaths")
ggplotly(l_indoor)

# High Outdoor Air Pollution
h_outdoor <- ggplot(high_poll, aes(x=country, y = avg_outdoor, color = avg_ozone, cex=5)) +
  geom_point() +
  labs(title = "Outdoor Air Pollution Deaths in High Population Countries") +
  xlab("Country") + 
  ylab("Average Deaths")
ggplotly(h_outdoor)

#Low Outdoor Air Pollution
l_outdoor <- ggplot(low_poll, aes(x = country, y= avg_outdoor, color = avg_ozone, cex=5)) +
  geom_point() +
  labs(title = "Outdoor Air Pollution Deaths in Low Population Countries")+
  xlab("Country")+
  ylab("Average Deaths")
ggplotly(l_outdoor)
```

***

# Pollution Over Time 
> Let's look at the previous two decades and compare the death count
> Has there been a change?

This is the first decade 1996-2006
```{r, echo=FALSE, message=FALSE}
# Look at the trend of deaths within the last decade from low and high populated countries
death_96_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 1996, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_96 = mean(total_deaths))

death_96_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 1996, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_96 = mean(total_deaths))

death_01_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2001, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_01 = mean(total_deaths))

death_01_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2001, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_01 = mean(total_deaths))

death_06_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2006, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_06 = mean(total_deaths))

death_06_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2006, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_06 = mean(total_deaths))

#cohesive table so it's easier to visualize 
#Low Population Table 2007-2017
all_low_first <- death_96_low %>% 
  mutate(death_01_low) %>% 
  mutate(death_06_low)

#High Population Table 2007-2017
all_high_first <- death_96_high %>% 
  mutate(death_01_high) %>% 
  mutate(death_06_high)

kable(list(all_high_first))
kable(list(all_low_first))
```
***

This is the second decade 2007-2017
```{r, echo=FALSE, message=FALSE}
# Look at the trend of deaths within the last decade from low and high populated countries
death_07_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2007, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_07 = mean(total_deaths))

death_07_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2007, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_07 = mean(total_deaths))

death_12_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2012, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_12 = mean(total_deaths))

death_12_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2012, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_12 = mean(total_deaths))

death_17_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2017, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_17 = mean(total_deaths))

death_17_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2017, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_17 = mean(total_deaths))

#cohesive table so it's easier to visualize 
#Low Population Table 2007-2017
all_low_second <- death_07_low %>% 
  mutate(death_12_low) %>% 
  mutate(death_17_low)

#High Population Table 2007-2017
all_high_second <- death_07_high %>% 
  mutate(death_12_high) %>% 
  mutate(death_17_high)

kable(list(all_high_second))
kable(list(all_low_second))
```

***

Let's graph the previous tables!

The first decade.
```{r, echo=FALSE}
#High Population Deaths 1996-2006 
p_high_first <- all_high_first %>%
  group_by(country) %>% 
  ggplot(aes(x= High_Deaths_96, y= High_Deaths_01, color= High_Deaths_06)) +
  geom_point() +
  facet_wrap(~country)

interact_high_first<- p_high_first + labs(title="High Populated Countries", x="1996 Deaths", y="2001 Deaths")

ggplotly(interact_high_first)

#Low Population Deaths 1996-2006
p_low_first <- all_low_first %>%
  group_by(country) %>% 
  ggplot(aes(x= Low_Deaths_96, y= Low_Deaths_01, color= Low_Deaths_06)) +
  geom_point() +
  facet_wrap(~country)

interact_low_first<- p_low_first + labs(title="Low Populated Countries", x="1996 Deaths", y="2001 Deaths")

ggplotly(interact_low_first)
```

***

This shows the second decade.
```{r, echo=FALSE}
#High Population Deaths 2007-2017
p_high_second <- all_high_second %>%
  group_by(country) %>% 
  ggplot(aes(x= High_Deaths_07, y= High_Deaths_12, color= High_Deaths_17)) +
  geom_point() +
  facet_wrap(~country)

interact_high_second<- p_high_second + labs(title="High Populated Countries", x="2007 Deaths", y="2012 Deaths")

ggplotly(interact_high_second)

#Low Population Deaths 2007-2017
p_low_second <- all_low_second %>%
  group_by(country) %>% 
  ggplot(aes(x= Low_Deaths_07, y= Low_Deaths_12, color= Low_Deaths_17)) +
  geom_point() +
  facet_wrap(~country)

interact_low_second<- p_low_second + labs(title="Low Populated Countries", x="2007 Deaths", y="2012 Deaths")

ggplotly(interact_low_second)

```

***

> By comparing each pollutant type, we can determine which year and country had the highest numbers of deaths

Indoor Deaths
```{r, echo=FALSE}
#Indoor 1996-2017
indoor_high<- deaths_df %>% 
    filter(year> 1995 & country %in% c('United States', 'Brazil','Nigeria', 
                                       'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country, year, indoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, indoor_deaths, color = country))+
  ggtitle("High Populated Indoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(indoor_high)

indoor_low<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Canada', 'Chile', 'Malawi', 
                                       'Serbia', 'Sri Lanka', 'Tonga')) %>% 
  group_by(country, year, indoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, indoor_deaths, color = country))+
  ggtitle("Low Populated Indoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(indoor_low)

#Pakistan and Malawi were both mainly affected in 1996
indoor_top<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Pakistan', 'Malawi')) %>% 
  group_by(country, year, indoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, indoor_deaths, color = country))+
  ggtitle("Greatest Indoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(indoor_top)
```

***

Outdoor Deaths
```{r, echo= FALSE}
#Outdoor Particulate 1996-2017
outdoor_high<- deaths_df %>% 
    filter(year> 1995 & country %in% c('United States', 'Brazil','Nigeria', 
                                       'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country, year, outdoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, outdoor_deaths, color = country))+
  ggtitle("High Populated Outdoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(outdoor_high)

outdoor_low<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Canada', 'Chile', 'Malawi', 
                                       'Serbia', 'Sri Lanka', 'Tonga')) %>% 
  group_by(country, year, outdoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, outdoor_deaths, color = country))+
  ggtitle("Low Populated Outdoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(outdoor_low)

#Serbia and Pakistan were the top countries affected. 
#2011 Pakistan had the greatest amount of Outdoor Deaths
outdoor_top<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Pakistan', 'Serbia')) %>% 
  group_by(country, year, outdoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, outdoor_deaths, color = country))+
  ggtitle("Greatest Outdoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(outdoor_top)
```

***
Ozone Deaths 
```{r, echo=FALSE}
#Outdoor ozone 1996-2017
ozone_high<- deaths_df %>% 
    filter(year> 1995 & country %in% c('United States', 'Brazil','Nigeria', 
                                       'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country, year, ozone_deaths) %>% 
  ggplot()+
  geom_point(aes(year, ozone_deaths, color = country))+
  ggtitle("High Populated Ozone Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(ozone_high)

ozone_low<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Canada', 'Chile', 'Malawi', 
                                       'Serbia', 'Sri Lanka', 'Tonga')) %>% 
  group_by(country, year, ozone_deaths) %>% 
  ggplot()+
  geom_point(aes(year, ozone_deaths, color = country))+
  ggtitle("Low Populated Ozone Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(ozone_low)

#Malawi and Pakistan were the top countries affected 
#1997 had the greatest amount of ozone death in Pakistan
ozone_top<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Pakistan', 'Malawi')) %>% 
  group_by(country, year, ozone_deaths) %>% 
  ggplot()+
  geom_point(aes(year, ozone_deaths, color = country))+
  ggtitle("Greatest Ozone Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(ozone_top)
```

# Which is worse? 
> outdoor or indoor pollution?

Let's reintroduce a graph we looked at earlier. Instead this time we will combine the pollutant types together.
```{r, echo=FALSE}
#Plot High Population
h_pollutant<- ggplot(high_poll, aes(x=avg_indoor, y = avg_outdoor, color= avg_ozone)) + 
  geom_point() +
  ggtitle("High Population Pollutant Type Death Counts")
ggplotly(h_pollutant)

#Plot Low Population
l_pollutant<- ggplot(low_poll, aes(x=avg_indoor, y = avg_outdoor, color= avg_ozone)) + 
  geom_point() +
  ggtitle("Low Population Pollutant Type Death Counts")
ggplotly(l_pollutant)
```

We cannot conclude which is worse.

* Low Populated Countries: 
  + it's half-and-half

* High Populated Countries: 
  + Outdoor pollution seems to be more detrimental with the exception of two countries in this sample set.

# We have this included already

```{r}
#Mean total deaths from 1996-2017 of high-population countries
deaths_highpop_countries <- deaths_df %>% 
  filter(country %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country) %>% 
  select(total_deaths) %>% 
  summarize(average_death_high = mean(total_deaths))

#Mean total deaths from 1990-2017 of high-population countries
deaths_lowpop_countries<- deaths_df %>% 
  filter(country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand')) %>% 
  group_by(country) %>% 
  select(total_deaths) %>% 
  summarize(average_death_low = mean(total_deaths))

#Average Death of both High and Low Populated countries  
kable(list(deaths_highpop_countries, deaths_lowpop_countries))

ggplot(deaths_highpop_countries)+
  geom_col(mapping = aes(x=country, y=average_death_high))+
             xlab("Country")+
             ylab("Average deaths (per 100,000)")+
             ggtitle("Average total deaths in high-population countries")+
  coord_flip()

ggplot(deaths_lowpop_countries)+
  geom_col(mapping = aes(x=country, y=average_death_low))+
             xlab("Country")+
             ylab("Average deaths (per 100,000)")+
             ggtitle("Average total deaths in low-population countries")+
  coord_flip()
```

# Summary

# Sources