### 2022-05-10 Final Report.Rmd
more explanations

---
title: "Final Report"
author: "Alison Cheek & Elizabeth Bekele"
date: '2022-05-06'
output: html_document
---
In our initial discussion over choosing a topic for our project, we narrowed it down to environmental-related data because we were interested in seeing possible trends over time and the vast quantity of environmental data is available.

# Introduction 
* Air pollution can be detrimental to both our health and the climate
  + Outdoor and indoor air pollution cause chronic pain, respiratory diseases, shortened lifespan
  + Air pollution kills about 7 million people worldwide every year 
  + Hopefully this information can showcase the importance of air pollution and that we should be more mindful about our planet

* Overview
  + We will see how different air pollution types affect the population 
  + compare past and present population numbers
  + determine which air pollutant type has the highest associated death rate

# Packages Required 
```{r, message=FALSE, warning=FALSE}
#This will allow us to filter through our data 
library(tidyverse)
library(dplyr)
#This will help us plot figures to showcase our findings
library(ggplot2)
#This will help us organize and display our data as necessary 
library(knitr)
library(kableExtra)
#This expands our plot uses 
library(plotly)
#Scientific Notation Disabled 
options(scipen=T)
```

# Deaths Data
We were excited to do our report over this data because it was relatively tidy and had quite a few categorical variables and options for additional columns to graph.

* Deaths from pollution:
  + Kaggle: [https://www.kaggle.com/datasets/akshat0giri/death-due-to-air-pollution-19902017] 
  + Author: Akshat Giri
  + Last update: 2 years ago

Our deaths due to air pollution data set was from Kaggle. The author is Akshat Giri and was last updated 2 years ago so it’s pretty relevant. When we first loaded in the data some of the column names were lengthy so we shortened them to: country, acronym, year, total deaths, indoor deaths, outdoor deaths, and ozone deaths.

Import the deaths-due-to-air-pollution data
```{r}
deaths_df_old <- data.frame(read.csv("death-rates-from-air-pollution.csv"))
glimpse(deaths_df_old)
```

We are going to rename a few of the columns and glimpse the data
```{r}
deaths_df<- deaths_df_old %>% rename(country=Entity, acronym=Code, year=Year, total_deaths=Air.pollution..total...deaths.per.100.000., indoor_deaths=Indoor.air.pollution..deaths.per.100.000., outdoor_deaths=Outdoor.particulate.matter..deaths.per.100.000., ozone_deaths=Outdoor.ozone.pollution..deaths.per.100.000.)

glimpse(deaths_df)
```

# Data Variables
Variables that interest us here include:

* __country__ 
* __total_deaths__: per 100,000
* __indoor_deaths__: Indoor air pollution is considered pollution that occurs in the household. Cooking with solid fuels:
  + Wood
  + Crop waste, dung
  + Charcoal, coal
  This method of cooking is commonly seen in underdeveloped countries.
* __outdoor_deaths__: Outdoor air or ambient air are emissions caused by combustion processes from motor vehicles, solid-fuel burning and industries
  + Ozone (O3)
  + Particulate matter (PM10 and PM2.5)
  + Nitrogen dioxide (NO2)
  + Carbon monoxide (CO)
  + Sulfur dioxide (SO2)

The data set takes a closer look at deaths caused by the ozone itself which is considered a component of outdoor air pollution.

* __ozone_deaths__: Ozone is a gas that occurs both in Earth's upper atmosphere and at ground level. Ozone in the atmosphere is an important and helpful greenhouse gas, but ground-level ozone is created by extensive use of fossil fuels: 
  + Pollutants emitted by cars
  + Power plants, industrial boilers, refineries, chemical plants

# World Population Data 
* World Population:
  + Kaggle: [https://www.kaggle.com/datasets/imdevskp/world-population-19602018] 
  + Author: Devakumar K. P. 
  + Last update: 2 Years ago

The world population data set is also from Kaggel. The author is Devakumar K. P. and was last updated 2 years ago so it is also recent. From looking at a glimpse of the data set you can see the columns are country name, year, and count which refers to the population at that time

Now, let's take a look at the population data.
```{r}
world_pop <- read.csv("population_total_long.csv")
glimpse(world_pop)
```

***

To get a general idea of 'deaths-dataframe' we made, let's make a plots to see what's happening. This is a plot of indoor x outdoor deaths around the world by country.

```{r, echo=FALSE}
d <- ggplot(deaths_df, aes(x = indoor_deaths, y = outdoor_deaths, text = paste0(country, ", ", year) )) + geom_point() +
  ggtitle("Outdoor Deaths vs Indoor Deaths")
ggplotly(d)
```

This is a mess, and so we chose two countries from each continent (a high-population and a low-population country) to graph.

***
We selected a high-population and a low-population country from each continent, but we wanted a consistent variation between our selection of low and high population countries. So, we came up with a formula for calculating what the low-population country should be by multiplying the high-population country by .10. For example, when we chose the U.S (which had a population of 331002651 at the time the data was recorded), we multiplied this number by .10 to get 33100265.1 and find the country whose population most-closesly matched (in this case it was Canada with 37742154).

We purposefully left out countries whose population numbers were higher than the majority because we didn't want those countries to skew the data (Russia, India, and China).

```{r, echo=FALSE}
#selecting high-population countries from the world population data frame 
high_pop_countries <- world_pop %>% 
  filter(Year> 1995 & Country.Name %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(Year)

high<- ggplot(high_pop_countries, aes(x=Country.Name, y=Count, color=Country.Name))+
  geom_boxplot()+
  labs(title="Selection of high-population countries", x="Country", y="Population (in millions)")
ggplotly(high)

#selecting low-population countries from the world population data frame 
low_pop_countries <- world_pop %>% 
  filter(Year> 1995 & Country.Name %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand'))%>% 
  group_by(Year)

low<- ggplot(low_pop_countries, aes(x=Country.Name, y=Count, color=Country.Name))+
  geom_boxplot()+
  labs(title="Selection of low-population countries", x="Country", y="Population (in millions)")
ggplotly(low)

kable(head(list(high_pop_countries[1:6,], low_pop_countries[1:6,])))

```

Continents:

North America: U.S, Canada
South America: Brazil, Chile
Africa: Nigeria, Malawi
Europe: Germany, Serbia
Asia: Pakistan, Sri Lanka
Oceania: Australia, New Zealand

# Combine Data Sets

First let's look at a table of the high and low populated countries using the world population data set.

```{r, echo=FALSE}
kable(head(list(high_pop_countries[1:6,], low_pop_countries[1:6,])))
```
***

Next, we are going to see the death count for high and low populated countries using the deaths dataframe.

```{r, echo=FALSE}
#selecting high-population deaths from death dataframe 
high_pop_death <- deaths_df %>% 
  filter(year > 1995 & country %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(year)

#selecting low-population deaths from death dataframe 
low_pop_death <- deaths_df %>% 
  filter(year > 1995 & country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand')) %>% 
  group_by(year)

kable(head(list(high_pop_death[1:22,], low_pop_death[1:22,])))
```
***

Lastly, we will join the population and and deaths with its respected country.

```{r, message=FALSE, echo=FALSE}
#Combined High Population with Pollution Death
joined_high <- right_join(high_pop_death, high_pop_countries, by= c('country' = 'Country.Name', 'year' = 'Year'))

#Combined Low Population with Pollution Death
joined_low <-right_join(low_pop_death, low_pop_countries, by= c('country' = 'Country.Name', 'year' = 'Year'))

joined_select <- full_join(joined_high, joined_low)

kable(head(list(joined_high[1:22,], joined_low[1:22,])))
```

***

We also looked at how the data varied by continent.

```{r}
joined_all <- right_join(deaths_df, world_pop, by=c('country' = 'Country.Name', 'year' = 'Year'))
head(joined_all)

north_america <- joined_all %>% filter(country %in% c("United States", "Canada"))
head(na.omit(north_america))

south_america <- joined_all %>% filter(country %in% c("Brazil", "Chile"))
head(na.omit(south_america))

africa <- joined_all %>% filter(country %in% c("Nigeria", "Malawi"))
head(na.omit(africa))

europe <- joined_all %>% filter(country %in% c("Germany", "Serbia"))
head(na.omit(europe))

asia <- joined_all %>% filter(country %in% c("Pakistan", "Sri Lanka"))
head(na.omit(asia))

oceania <- joined_all %>% filter(country %in% c("Australia", "New Zealand"))
head(na.omit(oceania))

```

***

This is a closer view on the population growth over time in both the high and low populated countries that we selected.

```{r, echo=FALSE}
pop_high_growth<- ggplot(data=joined_high, aes(x=year, y=Count, color=country)) +
  geom_line() +
  labs(title="Population Over Time in High Populated Countries")+
  xlab("Year")+
  ylab("Population")
ggplotly(pop_high_growth)

pop_low_growth<- ggplot(data=joined_low, aes(x=year, y=Count, color=country)) +
  geom_line() +
  labs(title="Population Over Time in Low Populated Countries")+
  xlab("Year")+
  ylab("Population")
ggplotly(pop_low_growth)
```

This graph shows the population change over time. Something to note is Germany and Australia's line seem relatively flat, but a closer look will determine there is a gradual increase that is drastically slower than the other countries.

```{r, echo=FALSE}
ggplot(data=joined_high, aes(x=year, y=Count, color=country, size = total_deaths/Count)) +
  geom_line() +
  labs(title="Percentage of Death Count in High Population Countries", subtitle="Bar width is equal to percentage of Deaths")+
  xlab("Year")+
  ylab("Population")

ggplot(data=joined_low, aes(x=year, y=Count, color=country, size = total_deaths/Count)) +
  geom_line() +
  labs(title="Percentage of Death Count in Low Population Countries", subtitle="Bar width is equal to percentage of Deaths")+
  xlab("Year")+
  ylab("Population")
```
These graphs are of the same information but we have added the percentage of air-pollution related deaths as the width of the line to demonstrate visually if deaths increased or decreased over time. It is easier to see in the high-populated countries but air-pollution related deaths do decrease over time.

# Death Count
> Which country has the highest average death count?

Let's make a table depicting the high and low populated countries and their respected death count due to pollution.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Mean total deaths of high-population countries
deaths_highpop_countries <- deaths_df %>% 
  filter(country %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country) %>% 
  select(total_deaths) %>% 
  summarize(hp_average_death = mean(total_deaths))

#Mean total deaths of high-population countries
deaths_lowpop_countries<- deaths_df %>% 
  filter(country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand')) %>% 
  group_by(country) %>% 
  select(total_deaths) %>% 
  summarize(lp_average_death = mean(total_deaths))

#Side-by-side Table to view deaths 
kable(list(deaths_highpop_countries, deaths_lowpop_countries))

```
We wanted to take a closer look at the death count and see which country has the highest average death count. In the tables we made you can see that Pakistan had the highest average death count at 144.33 for the high populated countries. Malawi had the highest average death count at 147.77 for the low populated countries, which is higher than Pakistan.

> Let's see how this is different from continent to continent

```{r}
#Mean total deaths for each continent
deaths_north <- na.omit(north_america)  %>% 
  group_by(country) %>% 
  summarize(north_america_deaths = mean(total_deaths))

deaths_south <- na.omit(south_america)  %>% 
  group_by(country) %>% 
  summarize(south_america_deaths = mean(total_deaths))

deaths_africa <- na.omit(africa)  %>% 
  group_by(country) %>% 
  summarize(africa_deaths = mean(total_deaths))

deaths_europe <- na.omit(europe)  %>% 
  group_by(country) %>% 
  summarize(europe_deaths = mean(total_deaths))

deaths_asia <- na.omit(asia)  %>% 
  group_by(country) %>% 
  summarize(asia_deaths = mean(total_deaths))

deaths_oceania <- na.omit(oceania)  %>% 
  group_by(country) %>% 
  summarize(oceania_deaths = mean(total_deaths))

#Table to view continent deaths 
kable(deaths_north, caption = "North America Average Death Count")
kable(deaths_south, caption = "South America Average Death Count")
kable(deaths_africa, caption = "Africa Average Death Count")
kable(deaths_asia, caption = "Asia Average Death Count")
kable(deaths_europe, caption = "Europe Average Death Count")
kable(deaths_oceania, caption = "Oceania Average Death Count")

```

When we look at the average death count based on continent we can see that overall Oceania countries had the least amount of deaths. On average Australia had an average death count of roughly 17.8 and New Zealand had an average death count of 15.9. Whereas Africa countries had the most amount of deaths. On average Malawi had an average of 147.8 and Nigeria had an average of 112.3.

***

Here's a graph to clearly visualize the previous table

```{r, echo=FALSE, message=FALSE}
#Plot High Population Deaths (average)
h<- ggplot(deaths_highpop_countries)+
  geom_col(mapping = aes(x=country, y=hp_average_death, fill=country))+
             xlab("Country")+
             ylab("Average deaths (per 100,000)")+
             ggtitle("Average deaths in high-population countries")+
  coord_flip()

ggplotly(h)

#Plot of Low Population Deaths (average)
l <- ggplot(deaths_lowpop_countries)+
  geom_col(mapping = aes(x=country, y=lp_average_death, fill=country))+
             xlab("Country")+
             ylab("Average deaths (per 100,000)")+
             ggtitle("Average deaths in low-population countries")+
  coord_flip()

ggplotly(l)
```

To get a better visualization we created a bar graph of the average deaths in both the high and low populated countries. In this high-population graph you can see that Pakistan is at the highest and Australia is at the lowest. In the low-population graph you can see that Malawi is at the highest and New Zealand is at the lowest.

*** 
So we've looked at the deaths due to pollution, but what percentage of the population was affected?

In order to get rid of the leading zeros, and clean up the y-axis, we multiplied the 'percent_high' and 'percent_low' by 100,000 since the data was per 100,000 when calculating deaths.

```{r, echo=FALSE, message=FALSE}
hp_countries_population <- world_pop %>% 
  filter(Country.Name %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia'), Year > 1995) %>% 
  group_by(Country.Name) %>% 
  select(Count) %>% 
  summarize(average_population = mean(Count))

lp_countries_population <- world_pop %>% 
  filter(Country.Name %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand'), Year > 1995) %>% 
  group_by(Country.Name) %>% 
  select(Count) %>% 
  summarize(average_population = mean(Count))

#Population Average Table
kable(list(hp_countries_population, lp_countries_population))

#Percentage of those affected 
affected_high<- joined_high %>% 
  mutate(percent_high = (total_deaths/mean(Count))*100000)

percent_high<- affected_high %>% ggplot(aes(x = year, y = percent_high, color = country))+
  geom_point() +
  labs(title = "High Population Countries Affected by Air Pollution") +
  xlab("Year")+
  ylab("Percent (total_deaths/average pop)")
ggplotly(percent_high)

affected_low <- joined_low %>% 
  mutate(percent_low = (total_deaths/mean(Count))*100000)

percent_low <- affected_low %>% ggplot(aes(x = year, y = percent_low, color = country))+
  geom_point() +
  labs(title = "Low Populated Countries Affected by Air Pollution") +
  xlab("Year")+
  ylab("Percent (total_deaths/average pop)")
ggplotly(percent_low)
```

So now that we’ve looked at the deaths due to pollution we wanted to see what percentage of the population was actually affected by this. At the top we have a table depicting the average populations in both the high and low populated countries. You can see for the high populated countries that Pakistan is in the lead with 12.1% and for the low populated countries Malawi is in the lead with 130.9%.

# Pollution Types
> Which type of pollution has the greatest number of deaths?

```{r, echo=FALSE, message=FALSE}
#High Population Pollutant Averages
high_poll <- deaths_df %>% 
  group_by(country) %>% 
  filter(country %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  select(country, indoor_deaths, outdoor_deaths, ozone_deaths) %>% 
  summarize(avg_indoor = mean(indoor_deaths), avg_outdoor = mean(outdoor_deaths), avg_ozone = mean(ozone_deaths)) %>% 
  arrange(desc(avg_indoor),desc(avg_outdoor),desc(avg_ozone))

kable(high_poll)

#Low Population Pollutant Averages 
low_poll <- deaths_df %>% 
  group_by(country) %>% 
  filter(country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand')) %>% 
  select(country, indoor_deaths, outdoor_deaths, ozone_deaths) %>% 
  summarize(avg_indoor = mean(indoor_deaths), avg_outdoor = mean(outdoor_deaths), avg_ozone = mean(ozone_deaths))

kable(low_poll)
```

***
```{r, echo=FALSE, warning=FALSE}

#Plot High Population
# High Indoor Air Pollution
h_indoor<- ggplot(high_poll, aes(x=country, y = avg_indoor, cex=5)) + 
  geom_point() +
  ggtitle("Indoor Air Pollution Deaths in High Population Countries")+
  xlab("Country")+
  ylab("Average Deaths")
ggplotly(h_indoor)

#Low Indoor Air Pollution
l_indoor<- ggplot(low_poll, aes(x=country, y = avg_indoor, cex=5)) + 
  geom_point() +
  ggtitle("Indoor Air Pollution Deaths in Low Population Countries")+
  xlab("Country")+
  ylab("Average Deaths")
ggplotly(l_indoor)

# High Outdoor Air Pollution
h_outdoor <- ggplot(high_poll, aes(x=country, y = avg_outdoor, color = avg_ozone, cex=5)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "violet", aesthetics = "colour") +
  geom_point() +
  labs(title = "Outdoor Air Pollution Deaths in High Population Countries") +
  xlab("Country") + 
  ylab("Average Deaths")
ggplotly(h_outdoor)

#Low Outdoor Air Pollution
l_outdoor <- ggplot(low_poll, aes(x = country, y= avg_outdoor, color = avg_ozone, cex=5)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "violet", aesthetics = "colour") +
  geom_point() +
  labs(title = "Outdoor Air Pollution Deaths in Low Population Countries")+
  xlab("Country")+
  ylab("Average Deaths")
ggplotly(l_outdoor)
```

***

# Pollution Over Time 
> Let's look at the previous two decades and compare the death count

> has there been a change?

This is the first decade 1996-2006
```{r, echo=FALSE, message=FALSE}
# Look at the trend of deaths within the last decade from low and high populated countries
death_96_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 1996, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_96 = mean(total_deaths))

death_96_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 1996, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_96 = mean(total_deaths))

death_01_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2001, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_01 = mean(total_deaths))

death_01_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2001, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_01 = mean(total_deaths))

death_06_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2006, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_06 = mean(total_deaths))

death_06_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2006, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_06 = mean(total_deaths))

#cohesive table so it's easier to visualize 
#Low Population Table 2007-2017
all_low_first <- death_96_low %>% 
  mutate(death_01_low) %>% 
  mutate(death_06_low)

#High Population Table 2007-2017
all_high_first <- death_96_high %>% 
  mutate(death_01_high) %>% 
  mutate(death_06_high)

kable(list(all_high_first))
kable(list(all_low_first))
```
***

This is the second decade 2007-2017
```{r, echo=FALSE, message=FALSE}
# Look at the trend of deaths within the last decade from low and high populated countries
death_07_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2007, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_07 = mean(total_deaths))

death_07_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2007, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_07 = mean(total_deaths))

death_12_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2012, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_12 = mean(total_deaths))

death_12_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2012, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_12 = mean(total_deaths))

death_17_high<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2017, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(High_Deaths_17 = mean(total_deaths))

death_17_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2017, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_17 = mean(total_deaths))

#cohesive table so it's easier to visualize 
#Low Population Table 2007-2017
all_low_second <- death_07_low %>% 
  mutate(death_12_low) %>% 
  mutate(death_17_low)

#High Population Table 2007-2017
all_high_second <- death_07_high %>% 
  mutate(death_12_high) %>% 
  mutate(death_17_high)

kable(list(all_high_second))
kable(list(all_low_second))
```

***
Let's see if there is variation by continent.
Here are some tables for the first decade (1996-2006) and second decade (2007-2017) grouped by continent.

```{r}
#North America 1996-2006
north_96 <- na.omit(north_america)  %>% 
  group_by(country) %>% 
  filter(year == 1996) %>% 
  summarize(avg_deaths_96 = mean(total_deaths))

north_01 <- na.omit(north_america)  %>% 
  group_by(country) %>% 
  filter(year == 2001) %>% 
  summarize(avg_deaths_01 = mean(total_deaths))

north_06 <- na.omit(north_america)  %>% 
  group_by(country) %>% 
  filter(year == 2006) %>% 
  summarize(avg_deaths_06 = mean(total_deaths))

kable(list(north_96,north_01,north_06), caption = "North America Deaths 1996-2006")

# North America 2007-2017
north_07 <- na.omit(north_america)  %>% 
  group_by(country) %>% 
  filter(year == 2007) %>% 
  summarize(avg_deaths_07 = mean(total_deaths))

north_12 <- na.omit(north_america)  %>% 
  group_by(country) %>% 
  filter(year == 2012) %>% 
  summarize(avg_deaths_12 = mean(total_deaths))

north_17 <- na.omit(north_america)  %>% 
  group_by(country) %>% 
  filter(year == 2017) %>% 
  summarize(avg_deaths_17 = mean(total_deaths))

kable(list(north_07,north_12,north_17), caption = "North America Deaths 2007-2017")

#South America 1996-2006
south_96 <- na.omit(south_america)  %>% 
  group_by(country) %>% 
  filter(year == 1996) %>% 
  summarize(avg_deaths_96 = mean(total_deaths))

south_01 <- na.omit(south_america)  %>% 
  group_by(country) %>% 
  filter(year == 2001) %>% 
  summarize(avg_deaths_01 = mean(total_deaths))

south_06 <- na.omit(south_america)  %>% 
  group_by(country) %>% 
  filter(year == 2006) %>% 
  summarize(avg_deaths_06 = mean(total_deaths))

kable(list(south_96,south_01,south_06), caption = "South America Deaths 1996-2006")

# South America 2007-2017
south_07 <- na.omit(south_america)  %>% 
  group_by(country) %>% 
  filter(year == 2007) %>% 
  summarize(avg_deaths_07 = mean(total_deaths))

south_12 <- na.omit(south_america)  %>% 
  group_by(country) %>% 
  filter(year == 2012) %>% 
  summarize(avg_deaths_12 = mean(total_deaths))

south_17 <- na.omit(south_america)  %>% 
  group_by(country) %>% 
  filter(year == 2017) %>% 
  summarize(avg_deaths_17 = mean(total_deaths))

kable(list(south_07,south_12,south_17), caption = "South America Deaths 2007-2017")

# Africa 1996-2006
africa_96 <- na.omit(africa)  %>% 
  group_by(country) %>% 
  filter(year == 1996) %>% 
  summarize(avg_deaths_96 = mean(total_deaths))

africa_01 <- na.omit(africa)  %>% 
  group_by(country) %>% 
  filter(year == 2001) %>% 
  summarize(avg_deaths_01 = mean(total_deaths))

africa_06 <- na.omit(africa)  %>% 
  group_by(country) %>% 
  filter(year == 2006) %>% 
  summarize(avg_deaths_06 = mean(total_deaths))

kable(list(africa_96,africa_01,africa_06), caption = "Africa Deaths 1996-2006")

# Africa 2007-2017
africa_07 <- na.omit(africa)  %>% 
  group_by(country) %>% 
  filter(year == 2007) %>% 
  summarize(avg_deaths_07 = mean(total_deaths))

africa_12 <- na.omit(africa)  %>% 
  group_by(country) %>% 
  filter(year == 2012) %>% 
  summarize(avg_deaths_12 = mean(total_deaths))

africa_17 <- na.omit(africa)  %>% 
  group_by(country) %>% 
  filter(year == 2017) %>% 
  summarize(avg_deaths_17 = mean(total_deaths))

kable(list(africa_07,africa_12,africa_17), caption = "Africa Deaths 2007-2017")

#Europe 1996-2006
europe_96 <- na.omit(europe)  %>% 
  group_by(country) %>% 
  filter(year == 1996) %>% 
  summarize(avg_deaths_96 = mean(total_deaths))

europe_01 <- na.omit(europe)  %>% 
  group_by(country) %>% 
  filter(year == 2001) %>% 
  summarize(avg_deaths_01 = mean(total_deaths))

europe_06 <- na.omit(europe)  %>% 
  group_by(country) %>% 
  filter(year == 2006) %>% 
  summarize(avg_deaths_06 = mean(total_deaths))

kable(list(europe_96,europe_01,europe_06), caption = "Europe Deaths 1996-2006")

#Europe 2007-2017
europe_07 <- na.omit(europe)  %>% 
  group_by(country) %>% 
  filter(year == 2007) %>% 
  summarize(avg_deaths_07 = mean(total_deaths))

europe_12 <- na.omit(europe)  %>% 
  group_by(country) %>% 
  filter(year == 2012) %>% 
  summarize(avg_deaths_12 = mean(total_deaths))

europe_17 <- na.omit(europe)  %>% 
  group_by(country) %>% 
  filter(year == 2017) %>% 
  summarize(avg_deaths_17 = mean(total_deaths))

kable(list(europe_07,europe_12,europe_17), caption = "Europe Deaths 2007-2017")

#Asia 1996-2006
asia_96 <- na.omit(asia)  %>% 
  group_by(country) %>% 
  filter(year == 1996) %>% 
  summarize(avg_deaths_96 = mean(total_deaths))

asia_01 <- na.omit(asia)  %>% 
  group_by(country) %>% 
  filter(year == 2001) %>% 
  summarize(avg_deaths_01 = mean(total_deaths))

asia_06 <- na.omit(asia)  %>% 
  group_by(country) %>% 
  filter(year == 2006) %>% 
  summarize(avg_deaths_06 = mean(total_deaths))

kable(list(asia_96,asia_01,asia_06), caption = "Asia Deaths 1996-2006")

#Asia 2007-2017
asia_07 <- na.omit(asia)  %>% 
  group_by(country) %>% 
  filter(year == 2007) %>% 
  summarize(avg_deaths_07 = mean(total_deaths))

asia_12 <- na.omit(asia)  %>% 
  group_by(country) %>% 
  filter(year == 2012) %>% 
  summarize(avg_deaths_12 = mean(total_deaths))

asia_17 <- na.omit(asia)  %>% 
  group_by(country) %>% 
  filter(year == 2017) %>% 
  summarize(avg_deaths_17 = mean(total_deaths))

kable(list(asia_07,asia_12,asia_17), caption = "Asia Deaths 2007-2017")

#Oceania 1996-2006
oceania_96 <- na.omit(oceania)  %>% 
  group_by(country) %>% 
  filter(year == 1996) %>% 
  summarize(avg_deaths_96 = mean(total_deaths))

oceania_01 <- na.omit(oceania)  %>% 
  group_by(country) %>% 
  filter(year == 2001) %>% 
  summarize(avg_deaths_01 = mean(total_deaths))

oceania_06 <- na.omit(oceania)  %>% 
  group_by(country) %>% 
  filter(year == 2006) %>% 
  summarize(avg_deaths_06 = mean(total_deaths))

kable(list(oceania_96,oceania_01,oceania_06), caption = "Oceania Deaths 1996-2006")

#Oceania 2007-2017
oceania_07 <- na.omit(oceania)  %>% 
  group_by(country) %>% 
  filter(year == 2007) %>% 
  summarize(avg_deaths_07 = mean(total_deaths))

oceania_12 <- na.omit(oceania)  %>% 
  group_by(country) %>% 
  filter(year == 2012) %>% 
  summarize(avg_deaths_12 = mean(total_deaths))

oceania_17 <- na.omit(oceania)  %>% 
  group_by(country) %>% 
  filter(year == 2017) %>% 
  summarize(avg_deaths_17 = mean(total_deaths))

kable(list(oceania_07,oceania_12,oceania_17), caption = "Oceania Deaths 2007-2017")
```
***

Let's graph the previous tables!

The first decade 1996-2006.
```{r, echo=FALSE}
#High Population Deaths 1996-2006
p_high_first <- all_high_first %>%
  group_by(country) %>% 
  ggplot(aes(x= High_Deaths_96, y= High_Deaths_01, color= High_Deaths_06)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_point() +
  facet_wrap(~country)

interact_high_first<- p_high_first + labs(title="High Populated Countries", x="1996 Deaths", y="2001 Deaths")

ggplotly(interact_high_first)

#High Population Deaths 1996 
p_high_first <- all_high_first %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= High_Deaths_96, fill = country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position = 'dodge')

interact_high_first<- p_high_first + labs(title="High Populated Countries", x="1996 Deaths")

ggplotly(interact_high_first)

#High Population Deaths 2001
p_high_first <- all_high_first %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= High_Deaths_01, fill = country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position = 'dodge')

interact_high_first<- p_high_first + labs(title="High Populated Countries", x="2001 Deaths")

ggplotly(interact_high_first)

#High Population Deaths 2006
p_high_first <- all_high_first %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= High_Deaths_06, fill = country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position = 'dodge')

interact_high_first<- p_high_first + labs(title="High Populated Countries", x="2006 Deaths")

ggplotly(interact_high_first)

#Low Population Deaths 1996-2006
p_low_first <- all_low_first %>%
  group_by(country) %>% 
  ggplot(aes(x= Low_Deaths_96, y= Low_Deaths_01, color= Low_Deaths_06)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_point() +
  facet_wrap(~country)

interact_low_first<- p_low_first + labs(title="Low Populated Countries", x="1996 Deaths", y="2001 Deaths")

ggplotly(interact_low_first)

#Low Population Deaths 1996 
p_low_first <- all_low_first %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= Low_Deaths_96, fill = country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position = 'dodge')

interact_low_first<- p_low_first + labs(title="High Populated Countries", x="1996 Deaths")

ggplotly(interact_low_first)

#Low Population Deaths 2001 
p_low_first <- all_low_first %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= Low_Deaths_01, fill = country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position = 'dodge')

interact_low_first<- p_low_first + labs(title="High Populated Countries", x="1996 Deaths")

ggplotly(interact_low_first)

#Low Population Deaths 2006 
p_low_first <- all_low_first %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= Low_Deaths_06, fill = country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position = 'dodge')

interact_low_first<- p_low_first + labs(title="High Populated Countries", x="1996 Deaths")

ggplotly(interact_low_first)
```

***

This shows the second decade 2007-2017.
```{r, echo=FALSE}
#High Population Deaths 2007-2017
p_high_second <- all_high_second %>%
  group_by(country) %>% 
  ggplot(aes(x= High_Deaths_07, y= High_Deaths_12, color= High_Deaths_17)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_point() +
  facet_wrap(~country)

interact_high_second<- p_high_second + labs(title="High Populated Countries", x="2007 Deaths", y="2017 Deaths")

ggplotly(interact_high_second)

#High Population Deaths 2007
p_high_second_07 <- all_high_second %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= High_Deaths_07, fill=country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position = 'dodge')

interact_high_second_07<- p_high_second_07 + labs(title="High Populated Countries", x="2007 Deaths")

ggplotly(interact_high_second_07)

#High Population Deaths 2012
p_high_second_12 <- all_high_second %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= High_Deaths_12, fill=country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position = 'dodge')

interact_high_second_12<- p_high_second_12 + labs(title="High Populated Countries", x="2012 Deaths")

ggplotly(interact_high_second_12)

#High Population Deaths 2017
p_high_second_17 <- all_high_second %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= High_Deaths_17, fill=country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position = 'dodge')

interact_high_second_17<- p_high_second_17 + labs(title="High Populated Countries", x="2017 Deaths")

ggplotly(interact_high_second_17)

#Low Population Deaths 2007-2017
p_low_second <- all_low_second %>%
  group_by(country) %>% 
  ggplot(aes(x= Low_Deaths_07, y= Low_Deaths_12, color= Low_Deaths_17)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_point() +
  facet_wrap(~country)

interact_low_second<- p_low_second + labs(title="Low Populated Countries", x="2007 Deaths", y="2017 Deaths")

ggplotly(interact_low_second)

#Low Population Deaths 2007
p_low_second_07 <- all_low_second %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= Low_Deaths_07, fill=country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position='dodge')

interact_low_second_07<- p_low_second_07 + labs(title="Low Populated Countries", x="2007 Deaths")

ggplotly(interact_low_second_07)

#Low Population Deaths 2012
p_low_second_12 <- all_low_second %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= Low_Deaths_12, fill=country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position='dodge')

interact_low_second_12<- p_low_second_12 + labs(title="Low Populated Countries", x="2012 Deaths")

ggplotly(interact_low_second_12)

#Low Population Deaths 2017
p_low_second_17 <- all_low_second %>%
  group_by(country) %>% 
  ggplot(aes(x= country, y= Low_Deaths_17, fill=country)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  geom_col(position='dodge')

interact_low_second_17<- p_low_second_17 + labs(title="Low Populated Countries", x="2017 Deaths")

ggplotly(interact_low_second_17)

```

***

> By comparing each pollutant type, we can determine which year and country had the highest numbers of deaths

Indoor Deaths
```{r, echo=FALSE}
#Indoor 1997-2017
indoor_high<- deaths_df %>% 
    filter(year> 1995 & country %in% c('United States', 'Brazil','Nigeria', 
                                       'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country, year, indoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, indoor_deaths, color = country))+
  ggtitle("High Populated Indoor Death 1997-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(indoor_high)

indoor_low<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Canada', 'Chile', 'Malawi', 
                                       'Serbia', 'Sri Lanka', 'Tonga')) %>% 
  group_by(country, year, indoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, indoor_deaths, color = country))+
  ggtitle("Low Populated Indoor Death 1997-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(indoor_low)

#Pakistan and Malawi were both mainly affected in 1996
indoor_top<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Pakistan', 'Malawi')) %>% 
  group_by(country, year, indoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, indoor_deaths, color = country))+
  ggtitle("Greatest Indoor Death 1997-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(indoor_top)
```

***

Outdoor Deaths
```{r, echo= FALSE}
#Outdoor Particulate 1997-2017
outdoor_high<- deaths_df %>% 
    filter(year> 1995 & country %in% c('United States', 'Brazil','Nigeria', 
                                       'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country, year, outdoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, outdoor_deaths, color = country))+
  ggtitle("High Populated Outdoor Death 1997-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(outdoor_high)

outdoor_low<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Canada', 'Chile', 'Malawi', 
                                       'Serbia', 'Sri Lanka', 'Tonga')) %>% 
  group_by(country, year, outdoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, outdoor_deaths, color = country))+
  ggtitle("Low Populated Outdoor Death 1997-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(outdoor_low)

#Serbia and Pakistan were the top countries affected. 
#2011 Pakistan had the greatest amount of Outdoor Deaths
outdoor_top<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Pakistan', 'Serbia')) %>% 
  group_by(country, year, outdoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, outdoor_deaths, color = country))+
  ggtitle("Greatest Outdoor Death 1997-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(outdoor_top)
```

***
Ozone Deaths 
```{r, echo=FALSE}
#Outdoor ozone 1997-2017
ozone_high<- deaths_df %>% 
    filter(year> 1995 & country %in% c('United States', 'Brazil','Nigeria', 
                                       'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country, year, ozone_deaths) %>% 
  ggplot()+
  geom_point(aes(year, ozone_deaths, color = country))+
  ggtitle("High Populated Ozone Death 1997-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(ozone_high)

ozone_low<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Canada', 'Chile', 'Malawi', 
                                       'Serbia', 'Sri Lanka', 'Tonga')) %>% 
  group_by(country, year, ozone_deaths) %>% 
  ggplot()+
  geom_point(aes(year, ozone_deaths, color = country))+
  ggtitle("Low Populated Ozone Death 1997-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(ozone_low)

#Malawi and Pakistan were the top countries affected 
#1997 had the greatest amount of ozone death in Pakistan
ozone_top<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Pakistan', 'Malawi')) %>% 
  group_by(country, year, ozone_deaths) %>% 
  ggplot()+
  geom_point(aes(year, ozone_deaths, color = country))+
  ggtitle("Greatest Ozone Death 1997-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(ozone_top)
```

# Which is worse? 
> outdoor or indoor pollution?

Let's reintroduce a graph we looked at earlier. Instead this time we will combine the pollutant types together.
```{r, echo=FALSE}
#Plot High Population
h_pollutant<- ggplot(high_poll, aes(x=avg_indoor, y = avg_outdoor, color= avg_ozone)) + 
  geom_point() +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  ggtitle("High Population Pollutant Type Death Counts")
ggplotly(h_pollutant)

#Plot Low Population
l_pollutant<- ggplot(low_poll, aes(x=avg_indoor, y = avg_outdoor, color= avg_ozone)) + 
  geom_point() +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "dark violet", aesthetics = "color") +
  ggtitle("Low Population Pollutant Type Death Counts")
ggplotly(l_pollutant)
```

We cannot conclude which is worse.

* High Populated Countries: 
  + Outdoor pollution seems to be more detrimental with the exception of two countries in this sample set.

* Low Populated Countries: 
  + it's half-and-half

# Summary

* Which country has the highest average death count?
  + High Population: Pakistan
  + Low Population: Malawi
* Has the percentage of the affected population decreased or increased over time?
  + Generally it is decreasing for both High and Low populated countries 
* Which pollutant type has the greatest number of deaths?
  + High Population: Indoor Pollution
  + Low Population: Indoor Pollution
* How has the death count changed over the past two decades?
  + 1996-2006: 
    - High Population: Decreases
    - Low Population: Decreases
  + 2007-2017: 
    - High Population: Decreases 
    - Low Population: Decreases
* Which year and country had the highest number of deaths per pollutant type?
  + We looked at years 1996-2017 
  + Indoor: Pakistan and Malawi were mainly affected in 1996
  + Outdoor: Serbia and Pakistan were the top countries. 
    - 2011 was the worst for Pakistan. 
    - 1997 was the worst for Serbia 
    - Sri Lanka and Tonga increased, but Sri Lanka had a steep decrease after 2015 
  + Outdoor Ozone: Pakistan and Malawi were the top countries. 
    - 1997 was the worst for Pakistan. 
    - 1998 was the worst for Malawi
    - United States was the second highest amount of deaths among the higher populated countries
    - Pakistan decreased and then slightly increased
* Which is worse - outdoor or indoor pollution?
  + Inconclusive

# Sources
* [https://www.kaggle.com/datasets/akshat0giri/death-due-to-air-pollution-19902017 ]

* [https://www.worldometers.info/population/world/]

* [https://www.who.int/health-topics/air-pollution#tab=tab_1]

* [https://www.epa.gov/ground-level-ozone-pollution/ground-level-ozone-basics]

* [https://www.health.nsw.gov.au/environment/air/Pages/outdoor-air-pollution.aspx]

* [https://www.mana.md/indoor-air-vs-outdoor-air/]

* [https://www.kaggle.com/datasets/imdevskp/world-population-19602018]